---
title: "EEMB 150 Quiz 5"
author: "Kerri Luttrell"
date: "3/1/2021"
output:
  pdf_document: default
  html_document: default
---
### *Anaerobic Oxidation of Methane*
 
**Set up the half reactions**

*oxidation half reaction*

 $CH_4 + 2H_2O \rightarrow CO_2 + 8H^+ +8e^-$

*reduction half reaction*

 $2e^- +2H^+  \rightarrow H_2$
 
 **Multiply reduction half reaction by 4**
 
 $8e^- +8H^+ \rightarrow 4H_2$
 
 **Cancel Out and Add**
 
  $8H^+ \rightarrow 4H_2$
  
  $CH_4 + 2H_2O \rightarrow CO_2 + 8H^+$

**Balanced Equation**

 $CH_4 + 2H_2O \rightarrow 4H_2+ CO_2$
 
```{r storing standard free energies}
H_pH0 <- 0
H_pH7 <- -39.83
H2 <- 0
sulfate <- -744.6
sulfide <- -27.87
CO2 <- -394.4
water <- -237.17
methane <- -50.75
deltaGred <-  (4*(H2)+(CO2)) -  (water*2 + methane)
deltaGred
```
### *Sulfate Reducing Bacteria*

**Set up the half reactions**

*oxidation half reaction*

 $H_2\rightarrow2e^- +2H^+$

*reduction half reaction*

 $SO_4^{2-} + 10H^+ + 8e^- \rightarrow H_2S +4H_2O$
 
 **Multiply oxidation half reaction by 4**
 
 $4H_2\rightarrow8e^- +8H^+$
 
 **Cancel Out and Add**
 
 $4H_2\rightarrow  8H^+$
  
 $SO_4^{2-} + 10H^+ \rightarrow H_2S +4H_2O$

**Balanced Equation**
 $SO_4^{2-} + 2H^+ +4H_2\rightarrow H_2S +4H_2O$
 
```{r}
 deltaGox <-  (4*(water)+(sulfide)) -  (H2*4 + sulfate + 2*H_pH0)
 deltaGox
```
 
 
 
 