---
title: "Thomas Fire Assignment"
author: "Kerri Luttrell"
date: "10/19/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(lubridate)
library(ggplot2)
library(psych)
library(RColorBrewer)
setwd("~/Github/144l_students/Input_Data/week1")

```

## Load Data
```{r data}
fire.data <- read_excel("~/Github/144l_students/Input_Data/week1/Thomas_Fire_Progression.xlsx")
```

```{r glimpse, include=FALSE}
glimpse(fire.data)
#vars= date, acres_burned, containment, PM10, PM25
pairs.panels(fire.data, lm=T, cor=T)

```

## Plot Data
```{r plot data}
p <- ggplot(fire.data, aes(x=Date))+ 
  geom_point(aes(y= PM10, colour = "PM10"))+
  geom_point(aes(y= Containment*1.5, colour = "Containment"))+
  scale_y_continuous(sec.axis = sec_axis(~./1.5, name = "Containment Level"))+
  scale_colour_manual(values = c("blue", "red"))+
  labs(y = "PM10 Abundance",
                x = "Date",
                colour = "Parameter", title = "PM10 Abundance and Containment Level over Time")+
  theme(legend.position = c(0.8, 0.6))
p
```

## Analysis 

 This data shows the relationship of containment levels and PM10 abundance over time. From this graph we can observe that PM10 abundance increased to nearly 150 the first few days of data collection. As the Thomas Fire became more contained, the PM10 levels eventually decreased and stayed at or below fifty. The oscillations below fifty could be due to this being a more ambient range subject to normal environmental dynamics rather than representative of particulate matter due to fire. This is important because it shows pretty rapid decline of PM10 abundance in response to fire containment, with no long lag times. This could inform recommendations on the duration of N95 face mask use.
